Stellar Radiation is Critical for Regulating Star Formation and Driving Outflows in Low Mass Dwarf Galaxies

### 2019-03-27 thesis/chapters/ch_chapter3.tex
updated section and figure labels with ch3 prepend to be consistent within thesis

We summarize our methods and physics models in Section~\ref{sec:methods}. We begin our analysis by presenting the only direct observable comparison we can make at this galaxy mass, discussing the metal retention fractions of our galaxy in Section~\ref{sec:ejection}. In Section~\ref{sec:log-normal} we focus on how each of our 15 individual metal species are distributed and evolve within each phase of the ISM. We discuss our results in Section~\ref{sec:discussion}, and conclude in Section~\ref{sec:conclusions}.

% AE: Benoit suggested subsections instead of paragraph splits... play with this to see what looks best
\section{Methods}
\label{sec:methods}
We refer the reader to Paper I for a detailed description of our numerical methods and feedback models. We briefly summarize the relevant details here.

@@

In this work we focus in detail on how each of the 15 individual metal species evolve in this galaxy. We address differences between the ejection fractions of each metal in Section~\ref{sec:ejection} and analyze for the first time the mass-fraction PDFs of the metals retained by the ISM in Section~\ref{sec:mixing}.

\subsection{Preferential Ejection of Metals from the ISM}
\label{sec:ejection}

\begin{figure*}
\centering
\includegraphics[width=0.99\linewidth]{figures/ch3/species_bar}\\
\includegraphics[width=0.99\linewidth]{figures/ch3/species_bar_ISM}
\caption{Top: The mass fraction of each metal species in the \textit{full simulation box} contained in the halo, gas in the galaxy disk, and stars. Bottom: The mass fraction of species within \textit{the disk} alone in each phase of the ISM (bottom). The leftmost bar in each plot shows the sum of all metals.}
\label{fig:species_fractions}
\end{figure*}

%mm [this all seems to have been deleted in one version, but doesn't look like it should have been.  Down to ">>>>"] <<<<<<< HEAD
The top panel of Figure~\ref{fig:species_fractions} gives the mass fraction of each element at the end of the simulation (500~Myr) in each of four reservoirs: locked in stars (yellow), in the ISM (blue), outside the galaxy but within the virial radius (purple), and outside the virial radius (salmon), including gas that has left the domain. We subdivide the ISM by phase in the bottom panel, giving the mass fractions of each element in the cold neutral medium (CNM, $f_{\rm H_2} < 0.5$,  T$< 100$~K), warm neutral medium (WNM, $10^2~\rm{K}\le \rm{T} < 10^4~\rm{K}$), warm ionized medium (WIM, $10^4~\rm{K}\le \rm{T} < 10^{5.5}~\rm{K}$), the hot ionized medium (HIM T$\ge 10^{5.5}$~K), and locked in stars (yellow).\footnote{The $f_{\rm H_2}$ restriction on the CNM is not relevant in our simulations, as there are no cells with $f_{\rm H_2} > 0.5$. $f_{\rm H_2}$ remains below about 0.35 for all cells (see Paper I).}
%mm Again, with the exception of  H and He, we [H & He aren't metals]
    We
only consider metals produced self-consistently through our star formation and stellar feedback methods; the initial mass of each metal species is zero.

@@

    Second,
nearly all of the metals 
%mm
    retained in the disk
reside within neutral gas, mostly in the CNM; only a few percent
reside in the hot phases. The exact fraction varies with each species, most notably for carbon; these fluctuations are at most $\sim 10$\%. This is not surprising, as the cold phases represent the majority of the mass in the ISM, but, as shown in Section~\ref{sec:statistics}, even though the cold phases harbor most of the metals, the hot phases have significantly higher metal mass fractions.

%The top panel of Figure~\ref{fig:species_fractions} gives the mass fraction of each element at the end of the simulation (500~Myr) in each of four reservoirs: locked in stars (yellow), in the ISM (blue), outside the galaxy but within the virial radius (purple), and outside the virial radius (salmon), including gas that has left the domain. We subdivide the ISM by phase in the bottom panel, giving the mass fractions of each element in the cold neutral medium (CNM, $f_{\rm H_2} < 0.5$,  T$< 100$~K), warm neutral medium (WNM, $10^2~\rm{K}\le \rm{T} < 10^4~\rm{K}$), warm ionized medium (WIM, $10^4~\rm{K}\le \rm{T} < 10^{5.5}~\rm{K}$), the hot ionized medium (HIM T$\ge 10^{5.5}$~K), and locked in stars (yellow). Again, with the exception of H and He, we only consider metals produced self-consistently through our star formation and stellar feedback methods; the initial mass of each metal species is zero.

%We find two major results. First, nearly all of the metals reside within neutral gas, mostly in the CNM; only a few percent reside in the hot phases. The exact fraction varies with each species, most notably for carbon; these fluctuations are at most $\sim 10$\%. This is not surprising, as the cold phases represent the majority of the mass in the ISM, but, as shown in Section~\ref{sec:statistics}, even though the cold phases harbor most of the metals, the hot phases have significantly higher metal mass fractions. Second, only a small fraction of produced metals are retained within the dwarf galaxy, in agreement with observations of nearby dwarf galaxies \citep[see][]{Kirby2011-metals, McQuinn2015}; however, the retention factor varies. The top panel shows a qualitative disagreement between the retention fractions of N and Ba (about 20\% for each) as compared to the rest of the metals ($\sim$ 4 -- 5\%). This suggests that individual metals \textit{do not} share the same dynamical evolution. Clearly, metal enrichment in galaxies is a phenomenon that cannot be fully captured using single, global metallicity field. In this particular case, assuming that all metals behave the same would underestimate the N and Ba enrichment by a factor of up to five, at least for low mass dwarf galaxies.
%mm [end of apparently deleted section] >>>>>>> 134a2fabac2aea2275cb3e8014308282b438742f

@@

\label{fig:species_sources}
\end{figure*}

The only physics that separates the dynamical evolution of these elements in our simulations is the individual sources of enrichment: AGB winds (stars less than 8~M$_{\odot}$), stellar winds (stars above 8~M$_{\odot}$), core collapse SNe, and Type Ia SNe. These channels differ in: 1) how long after a given star formation event they occur, 2), by consequence, the typical ISM properties in which they occur, and 3) how much energy is associated with each event, which determines ejecta temperature and velocity. To understand where each of the elements in Figure~\ref{fig:species_fractions} originated, we show the mass fraction of metals produced through each channel in Figure~\ref{fig:species_sources}. Core collapse SNe are responsible for over 90\% of the total metal enrichment in our galaxy, but this is clearly not true for all elements. In particular, a majority of N (74\%) and Ba (65\%) are ejected by AGB winds; both are retained at a higher rate than the rest of the metals. That this behavior exists for N and Ba in our simulations is dependent upon the metallicity of our galaxy and choice of stellar yield tables, but we can generalize this result to say that, for any assumed set of yields, low mass galaxies should more easily retain \textit{any} elements synthesized predominately in AGB winds, as compared to elements synthesized through SNe.\footnote{One might expect that much of the N and Ba that is ejected by galactic winds is comprised mostly of the $\sim$30\% of each species produced through the stellar winds of more massive stars and SNe. However, this cannot be proven as we lack Lagrangian information about individual gas elements.}
% AE: Actually if there are significant isotope variations between nucleosynthetic sources (i.e. N produced in AGB winds is a different isotope than the trace N produced in SNe) then that could be a very cool thing to distinguish if we had good gas-phase abundance measures of various isotopes... but maybe not if they aren't very stable....
We discuss how this result may extend towards other metal yields that are not well sampled on our relatively short (500 Myr) simulation timescales in Section~\ref{sec:discussion:metal yields}.

AGB winds have low energy and velocities (10~km~s$^{-1}$) as compared to the energy and typical expansion velocities of SNe ($\sim$1000~km~s$^{-1}$). In addition, their longer timescales relative to massive stars means that AGB stars are typically removed from their birth regions and randomly distributed through the galactic disk. The changes in typical ISM density and height of these events contributes to these variations. We show histograms of the average number density within 20~pc of any given enrichment source (top) and height above/below the disk (bottom) within 1~Myr of each event (as limited by our output cadence) in Figure~\ref{fig:spatial distribution}. As shown, SNe peak at very low densities, indicating that most explode in superbubbles, regions carved out by previous SNe. AGB stars predominately release their metals close to the average ISM density. The scale height distributions for both events show no significant differences. We do not expect these differences to be the dominant effect in driving the differential evolution of elements ejected by AGB winds vs. SNe, compared to the energetics, but can certainly play a significant role in determining the mixing behavior of individual enrichment events. The changes to mixing behavior as a function of ISM properties will be investigated in more detail in a future work.

\begin{figure}
\centering
\includegraphics[width=0.95\linewidth]{figures/ch3/SE_density}\\
\includegraphics[width=0.95\linewidth]{figures/ch3/SE_z}
\caption{The volume-averaged gas number densities within 20~pc of a given event (top) and vertical position above/below the disk (bottom) within 1~Myr before the event.}
\label{fig:spatial distribution}
\end{figure}

\subsection{Mixing and Distribution of Metals in the ISM}
\label{sec:mixing}
We characterize the metal distributions (in Sections~\ref{sec:log-normal} and \ref{sec:phase-pdfs}) and evolution (in Section~\ref{sec:statistics}) in our galaxy to build towards a  a complete model for how metals mix and evolve in the complex, multi-phase ISM of real galaxies.

\subsubsection{A Functional Form for Metal PDFs}
\label{sec:log-normal}

@@

The physics that drives the density PDF is directly related to the process of metal mixing and diffusion. However, there is no a priori reason why gas density and metallicity PDFs should have similar functional forms. We demonstrate here for the first time that the mass fraction PDFs for each metal species in our simulation can indeed be well fit using a piecewise log-normal + power-law PDF. We use a simple conceptual model in Section~\ref{sec:interpretation} to motivate the emergence of this distribution.

@@

We show in Figure~\ref{fig:log-normal} the numerical PDFs (solid histograms) and log-normal + power-law fits (dashed lines) across individual gas phases. These PDFs have been computed at an arbitrary single point in time in the middle of the simulation run.
% 270 Myr
For clarity, we only show a subset of the 15 elements we follow. As shown, there are clear differences in the PDFs across elements of different nucleosynthetic sources (AGB vs. SNe) and between each phase. However, each of these distributions are characterized by a power-law tail towards high metal fractions and a turnover of varying width at low metal fractions. We discuss the differences among each phase in more detail in the next section, but note here that the significance of each of these components varies notably across phases. In many of these cases, the piecewise log-normal + power-law distribution fits the numerical PDF quite well. However, there are often deviations, particularly at low metal fractions, from pure log-normal behavior. This manifests as either a very broad, flat PDF at low metal fraction (see N and Ba, for example), or large peaks not well described by $p(Z)$ at low metal fraction. In these situations, it is unclear what, if any, portion should be considered as a log-normal, or if there are multiple components within this region.

@@

\begin{figure*}
\centering
\includegraphics[width=0.95\linewidth]{figures/ch3/DD0390_element_by_element}
\caption{The numerical PDFs (solid histograms) and the associated log-normal + power-law fits (dashed lines) for a subset of the elements tracked in our simulation in each of the four gas phases defined in Section~\ref{sec:ejection}: CNM (dark blue), WNM (light blue), WIM (light orange), HIM (red), and all the gas in the ISM (black). For clarity, each distribution is normalized to the mode of the full-disk PDF (black) and is centered on the median value of the full-disk PDF. We note the vertical axis normalization is such that integrating over the shown PDF gives the mass fraction of that phase in the disk. Since the CNM dominates the mass fraction of our galaxy, the black curve is often obscured at low metal fractions.}
\label{fig:log-normal}
\end{figure*}

\subsubsection{PDF Variation Across Gas Phase}
\label{sec:phase-pdfs}

The various phases represented in Figure~\ref{fig:log-normal} involve variations in density and temperature of more than six orders of magnitude. The evolution of each phase is qualitatively different, and the metal mixing behavior of each should vary. Mixing timescales over a given length scale should be related to the local sound speed; hot gas, with higher sound speeds, should mix more rapidly than the dense, disconnected clumps of cold gas in the ISM. In addition, one would expect a metallicity gradient with gas temperature as enrichment occurs first in the hot phases, cooling and enriching denser gas over time. We examine the PDF variations among elements in each phase, as shown in Figure~\ref{fig:log-normal}, in more detail here.

@@

\subsubsection{The Time Evolution of Metal PDFs}
\label{sec:statistics}
We focus on the time evolution of the full numerical PDFs in this section. Our results here do not depend upon the choice of functional form for $p(Z)$. Figure~\ref{fig:phase-statistics} shows the evolution of four different statistics for the O (top) and Ba (bottom) PDFs. These two elements are treated as representative elements for SN and AGB wind production respectively. The difference between the mean and median masses of the PDF (second column) is a measure of the skew of the PDF. Positive values indicate that metals are preferentially sequestered in metal-rich gas, and are less well mixed throughout the given phase. The skew is always positive in these distributions.

\begin{figure*}
\centering
\includegraphics[width=0.95\linewidth]{figures/ch3/O_Ba_distribution_evolution}
\caption{Time evolution of three different statistics for the full distributions of O (top) and Ba (bottom) in each phase of our simulation. The panels show log$_{10}$ of the median (left), the difference, in dex, between the mean and the median (middle), and the 90$^{\rm th}$ decile and 10$^{\rm th}$ decile difference in dex (right).}
\label{fig:phase-statistics}
\end{figure*}

@@

Although these trends are generally true for Ba, its evolution is much more complicated.
%In addition, the HIM is much less offset from the rest of the phases than is seen in O, with much less variability.
Unlike O, the spread and skew in Ba for the CNM, WNM, and WIM \textit{increase} during the evolution, reaching differences in 90$^{\rm th}$ and 10$^{\rm th}$ percentiles of nearly 2 dex in the CNM and over 2 dex in the WNM and WIM. The HIM is seemingly unaffected by this trend, and simply fluctuates throughout the simulation. Given their lower energies, AGB winds more directly enrich the WNM and WIM, not the HIM as in SNe. The consequence of this is clear in Figure~\ref{fig:log-normal} by the wider PDFs and longer power-law tails in Ba in these phases. These tails represent the most recently enriched gas, 
%mm and 
        which
is clearly much more locally confined than O.

@@

To summarize, Figure~\ref{fig:phase-statistics} demonstrates: 1) there are qualitative differences in how SN-injected elements (e.g. O, Mg) and AGB-wind-injected elements (e.g. N, Ba) are distributed through the ISM, with the latter having a broader range of variation and being less well mixed in all phases except the HIM; 2) hotter phases are more metal enriched, both because the cooler phases make up most of the initially unenriched mass of the ISM and because the hotter phases are more directly populated by recent enrichment events; 3) the cooler, denser phases, particularly for SN injected elements, are more well mixed than the hot phases of the ISM; 4) the PDFs of metal mass fraction are best fit by a log-normal + power law function; and therefore 5) the median metallicity available for star formation 
%mm rates 
lies below the mean galactic value. In the case of SN injected elements, enrichment proceeds quickly through the HIM over the entire galaxy. For AGB injected elements, enrichment proceeds through the WIM and WNM, leading to longer mixing timescales and larger metal fraction variations in the ISM.

\section{Discussion}
\label{sec:discussion}
We begin with a simple toy model that motivates the power-law tail at high metal fractions of the metal fraction PDFs and a subsequent turnover at low metal fractions in Section~\ref{sec:interpretation}. This work is placed in context with previous papers focusing on metal mixing in the ISM in Section~\ref{sec:context}. We discuss the relevance of additional AGB yields not directly followed in this study in Section~\ref{sec:discussion:metal yields}. Finally, in Section~\ref{sec:stellar abundances} we discuss how these results relate to stellar abundances, make generalizations to more massive galaxies in Section~\ref{sec:massive galaxies}, and discuss possible impacts of these results on chemical enrichment from more exotic nucleosynthetic sources in Section~\ref{sec:exotic enrichment}.

\subsection{Physical Interpretation of the PDF}
\label{sec:interpretation}
%
% KVJ wanted clarification on this argument. I believe my added text 
% does this sufficiently.
Take the simple case of an initially primordial, uniform, isothermal medium of mass $M_o$ with initial metallicity $Z_{\rm o}$, containing a single, un-mixed enrichment event of mass $M_{\rm ej}$ whose size is small compared to the system and mass $M_{\rm ej} / M_o \ll 1$. Thus the background medium represents a virtually inexhaustible (but finite) source of un-enriched gas. In this case, $p(t,Z)$ initially takes the form of a double-delta function
\begin{equation}
p(t_{\rm o},Z) = \delta(Z_{\rm o}) + \frac{M_{\rm ej}}{M_o}\delta(Z - Z_{\rm ej}).
\end{equation}
%mm [edited the rest of this paragraph, and the next though I don't think there are any substantive changes unmarked.]
If the enriched material mixes continually with the un-enriched gas at a constant rate, after some time  $\tau_{\rm mix}$ this gas will have mixed with an equal amount of primordial gas. At this time, $p(\tau_{\rm mix},Z) = \delta(Z_{\rm o}) + (2M_{\rm ej}/M_o) \delta(Z - (1/2)Z_{\rm ej})$. Then, $p(2\tau_{\rm mix},Z) = \delta(Z_{\rm o}) + (4M_{\rm ej}/M_o)\delta(Z - (1/4) Z_{\rm ej})$, and so on for other multiples of $\tau_{\rm mix}$. This represents an inverse relationship between the mass of enriched gas and the metallicity of the enriched gas. The superposition of the distributions of a single event evolving in time will appear as a power-law distribution with slope $\alpha = 1$.

@@

\begin{figure}
\centering
\includegraphics[width=0.95\linewidth]{figures/ch3/stellar_CNM_distance}
\caption{The separation (in dex) of each star's oxygen fraction from the median value of the CNM oxygen mass fraction PDF for the time within 1~Myr (our time resolution) of each star's formation time. The median deviation is given in the plot for stars formed after the initial star formation and enrichment period (120~Myr.)}
\label{fig:stars}
\end{figure}

>>>>

We summarize our methods and physics models in Section~\ref{ch3:sec:methods}. We begin our analysis by presenting the only direct observable comparison we can make at this galaxy mass, discussing the metal retention fractions of our galaxy in Section~\ref{ch3:sec:ejection}. In Section~\ref{ch3:sec:log-normal} we focus on how each of our 15 individual metal species are distributed and evolve within each phase of the ISM. We discuss our results in Section~\ref{ch3:sec:discussion}, and conclude in Section~\ref{ch3:sec:conclusions}.

% AE: Benoit suggested subsections instead of paragraph splits... play with this to see what looks best
\section{Methods}
\label{ch3:sec:methods}
We refer the reader to Paper I for a detailed description of our numerical methods and feedback models. We briefly summarize the relevant details here.

@@

In this work we focus in detail on how each of the 15 individual metal species evolve in this galaxy. We address differences between the ejection fractions of each metal in Section~\ref{ch3:sec:ejection} and analyze for the first time the mass-fraction PDFs of the metals retained by the ISM in Section~\ref{ch3:sec:mixing}.

\subsection{Preferential Ejection of Metals from the ISM}
\label{ch3:sec:ejection}

\begin{figure*}
\centering
\includegraphics[width=0.99\linewidth]{figures/ch3/species_bar}\\
\includegraphics[width=0.99\linewidth]{figures/ch3/species_bar_ISM}
\caption{Top: The mass fraction of each metal species in the \textit{full simulation box} contained in the halo, gas in the galaxy disk, and stars. Bottom: The mass fraction of species within \textit{the disk} alone in each phase of the ISM (bottom). The leftmost bar in each plot shows the sum of all metals.}
\label{ch3:fig:species_fractions}
\end{figure*}

%mm [this all seems to have been deleted in one version, but doesn't look like it should have been.  Down to ">>>>"] <<<<<<< HEAD
The top panel of Figure~\ref{ch3:fig:species_fractions} gives the mass fraction of each element at the end of the simulation (500~Myr) in each of four reservoirs: locked in stars (yellow), in the ISM (blue), outside the galaxy but within the virial radius (purple), and outside the virial radius (salmon), including gas that has left the domain. We subdivide the ISM by phase in the bottom panel, giving the mass fractions of each element in the cold neutral medium (CNM, $f_{\rm H_2} < 0.5$,  T$< 100$~K), warm neutral medium (WNM, $10^2~\rm{K}\le \rm{T} < 10^4~\rm{K}$), warm ionized medium (WIM, $10^4~\rm{K}\le \rm{T} < 10^{5.5}~\rm{K}$), the hot ionized medium (HIM T$\ge 10^{5.5}$~K), and locked in stars (yellow).\footnote{The $f_{\rm H_2}$ restriction on the CNM is not relevant in our simulations, as there are no cells with $f_{\rm H_2} > 0.5$. $f_{\rm H_2}$ remains below about 0.35 for all cells (see Paper I).}
%mm Again, with the exception of  H and He, we [H & He aren't metals]
    We
only consider metals produced self-consistently through our star formation and stellar feedback methods; the initial mass of each metal species is zero.

@@

    Second,
nearly all of the metals 
%mm
    retained in the disk
reside within neutral gas, mostly in the CNM; only a few percent
reside in the hot phases. The exact fraction varies with each species, most notably for carbon; these fluctuations are at most $\sim 10$\%. This is not surprising, as the cold phases represent the majority of the mass in the ISM, but, as shown in Section~\ref{ch3:sec:statistics}, even though the cold phases harbor most of the metals, the hot phases have significantly higher metal mass fractions.

%The top panel of Figure~\ref{ch3:fig:species_fractions} gives the mass fraction of each element at the end of the simulation (500~Myr) in each of four reservoirs: locked in stars (yellow), in the ISM (blue), outside the galaxy but within the virial radius (purple), and outside the virial radius (salmon), including gas that has left the domain. We subdivide the ISM by phase in the bottom panel, giving the mass fractions of each element in the cold neutral medium (CNM, $f_{\rm H_2} < 0.5$,  T$< 100$~K), warm neutral medium (WNM, $10^2~\rm{K}\le \rm{T} < 10^4~\rm{K}$), warm ionized medium (WIM, $10^4~\rm{K}\le \rm{T} < 10^{5.5}~\rm{K}$), the hot ionized medium (HIM T$\ge 10^{5.5}$~K), and locked in stars (yellow). Again, with the exception of H and He, we only consider metals produced self-consistently through our star formation and stellar feedback methods; the initial mass of each metal species is zero.

%We find two major results. First, nearly all of the metals reside within neutral gas, mostly in the CNM; only a few percent reside in the hot phases. The exact fraction varies with each species, most notably for carbon; these fluctuations are at most $\sim 10$\%. This is not surprising, as the cold phases represent the majority of the mass in the ISM, but, as shown in Section~\ref{ch3:sec:statistics}, even though the cold phases harbor most of the metals, the hot phases have significantly higher metal mass fractions. Second, only a small fraction of produced metals are retained within the dwarf galaxy, in agreement with observations of nearby dwarf galaxies \citep[see][]{Kirby2011-metals, McQuinn2015}; however, the retention factor varies. The top panel shows a qualitative disagreement between the retention fractions of N and Ba (about 20\% for each) as compared to the rest of the metals ($\sim$ 4 -- 5\%). This suggests that individual metals \textit{do not} share the same dynamical evolution. Clearly, metal enrichment in galaxies is a phenomenon that cannot be fully captured using single, global metallicity field. In this particular case, assuming that all metals behave the same would underestimate the N and Ba enrichment by a factor of up to five, at least for low mass dwarf galaxies.
%mm [end of apparently deleted section] >>>>>>> 134a2fabac2aea2275cb3e8014308282b438742f

@@

\label{ch3:fig:species_sources}
\end{figure*}

The only physics that separates the dynamical evolution of these elements in our simulations is the individual sources of enrichment: AGB winds (stars less than 8~M$_{\odot}$), stellar winds (stars above 8~M$_{\odot}$), core collapse SNe, and Type Ia SNe. These channels differ in: 1) how long after a given star formation event they occur, 2), by consequence, the typical ISM properties in which they occur, and 3) how much energy is associated with each event, which determines ejecta temperature and velocity. To understand where each of the elements in Figure~\ref{ch3:fig:species_fractions} originated, we show the mass fraction of metals produced through each channel in Figure~\ref{ch3:fig:species_sources}. Core collapse SNe are responsible for over 90\% of the total metal enrichment in our galaxy, but this is clearly not true for all elements. In particular, a majority of N (74\%) and Ba (65\%) are ejected by AGB winds; both are retained at a higher rate than the rest of the metals. That this behavior exists for N and Ba in our simulations is dependent upon the metallicity of our galaxy and choice of stellar yield tables, but we can generalize this result to say that, for any assumed set of yields, low mass galaxies should more easily retain \textit{any} elements synthesized predominately in AGB winds, as compared to elements synthesized through SNe.\footnote{One might expect that much of the N and Ba that is ejected by galactic winds is comprised mostly of the $\sim$30\% of each species produced through the stellar winds of more massive stars and SNe. However, this cannot be proven as we lack Lagrangian information about individual gas elements.}
% AE: Actually if there are significant isotope variations between nucleosynthetic sources (i.e. N produced in AGB winds is a different isotope than the trace N produced in SNe) then that could be a very cool thing to distinguish if we had good gas-phase abundance measures of various isotopes... but maybe not if they aren't very stable....
We discuss how this result may extend towards other metal yields that are not well sampled on our relatively short (500 Myr) simulation timescales in Section~\ref{ch3:sec:discussion:metal yields}.

AGB winds have low energy and velocities (10~km~s$^{-1}$) as compared to the energy and typical expansion velocities of SNe ($\sim$1000~km~s$^{-1}$). In addition, their longer timescales relative to massive stars means that AGB stars are typically removed from their birth regions and randomly distributed through the galactic disk. The changes in typical ISM density and height of these events contributes to these variations. We show histograms of the average number density within 20~pc of any given enrichment source (top) and height above/below the disk (bottom) within 1~Myr of each event (as limited by our output cadence) in Figure~\ref{ch3:fig:spatial distribution}. As shown, SNe peak at very low densities, indicating that most explode in superbubbles, regions carved out by previous SNe. AGB stars predominately release their metals close to the average ISM density. The scale height distributions for both events show no significant differences. We do not expect these differences to be the dominant effect in driving the differential evolution of elements ejected by AGB winds vs. SNe, compared to the energetics, but can certainly play a significant role in determining the mixing behavior of individual enrichment events. The changes to mixing behavior as a function of ISM properties will be investigated in more detail in a future work.

\begin{figure}
\centering
\includegraphics[width=0.95\linewidth]{figures/ch3/SE_density}\\
\includegraphics[width=0.95\linewidth]{figures/ch3/SE_z}
\caption{The volume-averaged gas number densities within 20~pc of a given event (top) and vertical position above/below the disk (bottom) within 1~Myr before the event.}
\label{ch3:fig:spatial distribution}
\end{figure}

\subsection{Mixing and Distribution of Metals in the ISM}
\label{ch3:sec:mixing}
We characterize the metal distributions (in Sections~\ref{ch3:sec:log-normal} and \ref{ch3:sec:phase-pdfs}) and evolution (in Section~\ref{ch3:sec:statistics}) in our galaxy to build towards a  a complete model for how metals mix and evolve in the complex, multi-phase ISM of real galaxies.

\subsubsection{A Functional Form for Metal PDFs}
\label{ch3:sec:log-normal}

@@

The physics that drives the density PDF is directly related to the process of metal mixing and diffusion. However, there is no a priori reason why gas density and metallicity PDFs should have similar functional forms. We demonstrate here for the first time that the mass fraction PDFs for each metal species in our simulation can indeed be well fit using a piecewise log-normal + power-law PDF. We use a simple conceptual model in Section~\ref{ch3:sec:interpretation} to motivate the emergence of this distribution.

@@

We show in Figure~\ref{ch3:fig:log-normal} the numerical PDFs (solid histograms) and log-normal + power-law fits (dashed lines) across individual gas phases. These PDFs have been computed at an arbitrary single point in time in the middle of the simulation run.
% 270 Myr
For clarity, we only show a subset of the 15 elements we follow. As shown, there are clear differences in the PDFs across elements of different nucleosynthetic sources (AGB vs. SNe) and between each phase. However, each of these distributions are characterized by a power-law tail towards high metal fractions and a turnover of varying width at low metal fractions. We discuss the differences among each phase in more detail in the next section, but note here that the significance of each of these components varies notably across phases. In many of these cases, the piecewise log-normal + power-law distribution fits the numerical PDF quite well. However, there are often deviations, particularly at low metal fractions, from pure log-normal behavior. This manifests as either a very broad, flat PDF at low metal fraction (see N and Ba, for example), or large peaks not well described by $p(Z)$ at low metal fraction. In these situations, it is unclear what, if any, portion should be considered as a log-normal, or if there are multiple components within this region.

@@

\begin{figure*}
\centering
\includegraphics[width=0.95\linewidth]{figures/ch3/DD0390_element_by_element}
\caption{The numerical PDFs (solid histograms) and the associated log-normal + power-law fits (dashed lines) for a subset of the elements tracked in our simulation in each of the four gas phases defined in Section~\ref{ch3:sec:ejection}: CNM (dark blue), WNM (light blue), WIM (light orange), HIM (red), and all the gas in the ISM (black). For clarity, each distribution is normalized to the mode of the full-disk PDF (black) and is centered on the median value of the full-disk PDF. We note the vertical axis normalization is such that integrating over the shown PDF gives the mass fraction of that phase in the disk. Since the CNM dominates the mass fraction of our galaxy, the black curve is often obscured at low metal fractions.}
\label{ch3:fig:log-normal}
\end{figure*}

\subsubsection{PDF Variation Across Gas Phase}
\label{ch3:sec:phase-pdfs}

The various phases represented in Figure~\ref{ch3:fig:log-normal} involve variations in density and temperature of more than six orders of magnitude. The evolution of each phase is qualitatively different, and the metal mixing behavior of each should vary. Mixing timescales over a given length scale should be related to the local sound speed; hot gas, with higher sound speeds, should mix more rapidly than the dense, disconnected clumps of cold gas in the ISM. In addition, one would expect a metallicity gradient with gas temperature as enrichment occurs first in the hot phases, cooling and enriching denser gas over time. We examine the PDF variations among elements in each phase, as shown in Figure~\ref{ch3:fig:log-normal}, in more detail here.

@@

\subsubsection{The Time Evolution of Metal PDFs}
\label{ch3:sec:statistics}
We focus on the time evolution of the full numerical PDFs in this section. Our results here do not depend upon the choice of functional form for $p(Z)$. Figure~\ref{ch3:fig:phase-statistics} shows the evolution of four different statistics for the O (top) and Ba (bottom) PDFs. These two elements are treated as representative elements for SN and AGB wind production respectively. The difference between the mean and median masses of the PDF (second column) is a measure of the skew of the PDF. Positive values indicate that metals are preferentially sequestered in metal-rich gas, and are less well mixed throughout the given phase. The skew is always positive in these distributions.

\begin{figure*}
\centering
\includegraphics[width=0.95\linewidth]{figures/ch3/O_Ba_distribution_evolution}
\caption{Time evolution of three different statistics for the full distributions of O (top) and Ba (bottom) in each phase of our simulation. The panels show log$_{10}$ of the median (left), the difference, in dex, between the mean and the median (middle), and the 90$^{\rm th}$ decile and 10$^{\rm th}$ decile difference in dex (right).}
\label{ch3:fig:phase-statistics}
\end{figure*}

@@

Although these trends are generally true for Ba, its evolution is much more complicated.
%In addition, the HIM is much less offset from the rest of the phases than is seen in O, with much less variability.
Unlike O, the spread and skew in Ba for the CNM, WNM, and WIM \textit{increase} during the evolution, reaching differences in 90$^{\rm th}$ and 10$^{\rm th}$ percentiles of nearly 2 dex in the CNM and over 2 dex in the WNM and WIM. The HIM is seemingly unaffected by this trend, and simply fluctuates throughout the simulation. Given their lower energies, AGB winds more directly enrich the WNM and WIM, not the HIM as in SNe. The consequence of this is clear in Figure~\ref{ch3:fig:log-normal} by the wider PDFs and longer power-law tails in Ba in these phases. These tails represent the most recently enriched gas, 
%mm and 
        which
is clearly much more locally confined than O.

@@

To summarize, Figure~\ref{ch3:fig:phase-statistics} demonstrates: 1) there are qualitative differences in how SN-injected elements (e.g. O, Mg) and AGB-wind-injected elements (e.g. N, Ba) are distributed through the ISM, with the latter having a broader range of variation and being less well mixed in all phases except the HIM; 2) hotter phases are more metal enriched, both because the cooler phases make up most of the initially unenriched mass of the ISM and because the hotter phases are more directly populated by recent enrichment events; 3) the cooler, denser phases, particularly for SN injected elements, are more well mixed than the hot phases of the ISM; 4) the PDFs of metal mass fraction are best fit by a log-normal + power law function; and therefore 5) the median metallicity available for star formation 
%mm rates 
lies below the mean galactic value. In the case of SN injected elements, enrichment proceeds quickly through the HIM over the entire galaxy. For AGB injected elements, enrichment proceeds through the WIM and WNM, leading to longer mixing timescales and larger metal fraction variations in the ISM.

\section{Discussion}
\label{ch3:sec:discussion}
We begin with a simple toy model that motivates the power-law tail at high metal fractions of the metal fraction PDFs and a subsequent turnover at low metal fractions in Section~\ref{ch3:sec:interpretation}. This work is placed in context with previous papers focusing on metal mixing in the ISM in Section~\ref{ch3:sec:context}. We discuss the relevance of additional AGB yields not directly followed in this study in Section~\ref{ch3:sec:discussion:metal yields}. Finally, in Section~\ref{ch3:sec:stellar abundances} we discuss how these results relate to stellar abundances, make generalizations to more massive galaxies in Section~\ref{ch3:sec:massive galaxies}, and discuss possible impacts of these results on chemical enrichment from more exotic nucleosynthetic sources in Section~\ref{ch3:sec:exotic enrichment}.

\subsection{Physical Interpretation of the PDF}
\label{ch3:sec:interpretation}
%
% KVJ wanted clarification on this argument. I believe my added text 
% does this sufficiently.
Take the simple case of an initially primordial, uniform, isothermal medium of mass $M_o$ with initial metallicity $Z_{\rm o}$, containing a single, un-mixed enrichment event of mass $M_{\rm ej}$ whose size is small compared to the system and mass $M_{\rm ej} / M_o \ll 1$. Thus the background medium represents a virtually inexhaustible (but finite) source of un-enriched gas. In this case, $p(t,Z)$ initially takes the form of a double-delta function
\begin{equation}
p(t_{\rm o},Z) = \delta(Z_{\rm o}) + \frac{M_{\rm ej}}{M_o}\delta(Z - Z_{\rm ej}).
\end{equation}
%mm [edited the rest of this paragraph, and the next though I don't think there are any substantive changes unmarked.]
If the enriched material mixes continually with the un-enriched gas at a constant rate, after some time  $\tau_{\rm mix}$ this gas will have mixed with an equal amount of primordial gas. At this time, $p(\tau_{\rm mix},Z) = \delta(Z_{\rm o}) + (2M_{\rm ej}/M_o) \delta(Z - (1/2)Z_{\rm ej})$. Then, $p(2\tau_{\rm mix},Z) = \delta(Z_{\rm o}) + (4M_{\rm ej}/M_o)\delta(Z - (1/4) Z_{\rm ej})$, and so on for other multiples of $\tau_{\rm mix}$. This represents an inverse relationship between the mass of enriched gas and the metallicity of the enriched gas. The superposition of the distributions of a single event evolving in time will appear as a power-law distribution with slope $\alpha = 1$.

@@

\begin{figure}
\centering
\includegraphics[width=0.95\linewidth]{figures/ch3/stellar_CNM_distance}
\caption{The separation (in dex) of each star's oxygen fraction from the median value of the CNM oxygen mass fraction PDF for the time within 1~Myr (our time resolution) of each star's formation time. The median deviation is given in the plot for stars formed after the initial star formation and enrichment period (120~Myr.)}
\label{ch3:fig:stars}
\end{figure}
